\section{Principal Component Analysis (PCA)}

\subsection{1. Data Preprocessing}
\textbf{Task:} Present centered data.

\paragraph{Answer.}
After subtracting the mean of each feature from the original matrix \(B\), the centered data is:
\[
B_c=
\begin{bmatrix}
-1 & -3 & -4\\
-1 & -1 & -2\\
 1 &  1 &  2\\
 1 &  3 &  4
\end{bmatrix}.
\]
This centered matrix is the input used for covariance and eigen analysis.

\subsection{2. Covariance Structure}
\textbf{Task:} Present covariance/scatter matrix.

\paragraph{Answer.}
The covariance matrix is:
\[
\Sigma=
\begin{bmatrix}
1.33 & 2.67 & 4.00\\
2.67 & 6.67 & 9.33\\
4.00 & 9.33 & 13.33
\end{bmatrix}.
\]
The matrix is symmetric, as expected. Diagonal entries are the variances of each feature, while off-diagonal entries are pairwise covariances. The third feature has the largest variance and also strong covariance with the other features, which already suggests correlated (redundant) structure.

\subsection{3. Eigen Decomposition}
\textbf{Task:} Derive eigenvalues of covariance matrix.

\paragraph{Answer.}
The eigenvalues are approximately:
\[
\lambda = [21.08,\ 0.25,\ 0.00].
\]
One dominant eigenvalue and two very small ones indicate that most variance is concentrated in a single direction.

\subsection{4. Projection}
\textbf{Task:} Project data onto first principal component.

\paragraph{Answer.}
Dimensionality reduction is obtained by projecting each centered observation onto the eigenvector associated with the largest eigenvalue (the first principal component). This maps each 3D point to a 1D coordinate while preserving most of the variance.

For this dataset, the projected coordinates are approximately:
\[
[5.09,\ 2.39,\ -2.39,\ -5.09].
\]
These values represent each sample position along the main variance axis.

\subsection{5. Interpretation}
\textbf{Task:} Discuss redundancy, intrinsic dimensionality, and explained variance.

\paragraph{Answer.}
The eigenvalue distribution shows that nearly all variance is captured by the first principal component. Therefore:
\begin{itemize}
  \item the variables are highly redundant;
  \item the intrinsic dimensionality is much lower than 3 (close to 1);
  \item a 1D PCA projection keeps almost all useful information.
\end{itemize}
In practical terms, this dataset can be summarized accurately with one principal component, and the remaining components contribute little additional information.
