\section{Principal Component Analysis (PCA)}

\subsection{1. Data Preprocessing}
\textbf{Task:} Present centered data.

\paragraph{Answer.}
Given
\[
B=\begin{bmatrix}
2 & 2 & 4\\
2 & 4 & 6\\
4 & 6 & 10\\
4 & 8 & 12
\end{bmatrix},
\quad
\mu=\begin{bmatrix}3 & 5 & 8\end{bmatrix},
\]
the centered data matrix is
\[
B_c=B-\mu=
\begin{bmatrix}
-1 & -3 & -4\\
-1 & -1 & -2\\
 1 &  1 &  2\\
 1 &  3 &  4
\end{bmatrix}.
\]

\subsection{2. Covariance Structure}
\textbf{Task:} Present covariance/scatter matrix.

\paragraph{Answer.}
Using the centered matrix with variables in columns, the covariance matrix is:
\[
\Sigma=
\begin{bmatrix}
1.3333 & 2.6667 & 4.0000\\
2.6667 & 6.6667 & 9.3333\\
4.0000 & 9.3333 & 13.3333
\end{bmatrix}.
\]
All off-diagonal entries are positive and large, which indicates strong correlation between features and therefore substantial redundancy in the original representation.

\subsection{3. Eigen Decomposition}
\textbf{Task:} Derive eigenvalues of covariance matrix.

\paragraph{Answer.}
The eigenvalues (sorted in descending order) are:
\[
\lambda_1=21.0803,\quad
\lambda_2=0.2530,\quad
\lambda_3\approx 0.
\]
The near-zero third eigenvalue indicates that one direction contributes almost no additional variance, consistent with an almost lower-dimensional structure.

A corresponding first principal direction (unit eigenvector) is:
\[
v_1=
\begin{bmatrix}
-0.2365\\
-0.5586\\
-0.7950
\end{bmatrix}.
\]

\subsection{4. Projection}
\textbf{Task:} Project data onto first principal component.

\paragraph{Answer.}
Projection onto the first principal component is computed by:
\[
z_i = B_{c,i}\,v_1.
\]
The resulting 1D coordinates are:
\[
z=
\begin{bmatrix}
5.0923,\ 2.3851,\ -2.3851,\ -5.0923
\end{bmatrix}.
\]
This 1D representation preserves almost all information while greatly reducing dimensionality.

\subsection{5. Interpretation}
\textbf{Task:} Discuss redundancy, intrinsic dimensionality, and explained variance.

\paragraph{Answer.}
Explained variance ratios from the eigenvalues are approximately:
\[
\frac{\lambda_1}{\sum\lambda}=98.81\%,\quad
\frac{\lambda_2}{\sum\lambda}=1.19\%,\quad
\frac{\lambda_3}{\sum\lambda}\approx 0\%.
\]
Therefore:
\begin{itemize}
  \item most variance is concentrated in the first principal component;
  \item the effective intrinsic dimensionality is close to 1;
  \item the original three features are highly redundant.
\end{itemize}
In practice, projecting onto the first principal component is an accurate low-dimensional summary for this dataset.
